---
title: "ECO395M STAT LEARNING Homework 2" 
author: "Mingwei Li, Xinyu Leng, Hongjin Long"
thanks: "Mingwei Li, Xinyu Leng and Hongjin Long are master students of economics, The University of Texas at Austin"
output:
  pdf_document: 
    number_sections: yes
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\begin{abstract}
This document is the second homework of ECO395M STAT LEARNING. 
\begin{figure}[!htbp]
	\centering
	\includegraphics[width=0.85\textwidth]{pics/0.jpg}
\end{figure}
\end{abstract}

\newpage
\tableofcontents

\newpage
\section{Problem 1: visualization}
\subsection{line graphs}
```{r,echo=FALSE,message=FALSE}
# Load the data and package
library(ggplot2)
library(tidyverse)
capmetro = read.csv("data/capmetro.csv")
```
```{r,echo=FALSE,message=FALSE}
# Extract the month and year
tem = capmetro$timestamp[1]
month_func = function(my_str){
  return(as.numeric(strsplit(my_str, "-", fixed=TRUE)[[1]][2]))
}
hour_func = function(my_str){
detailed_time = strsplit(my_str, " ", fixed=TRUE)[[1]][2]
hour = as.numeric(strsplit(detailed_time, ":", fixed=TRUE)[[1]][1])
return(hour)
}
capmetro$month = apply(array(capmetro$timestamp),1,month_func)
capmetro$hour = apply(array(capmetro$timestamp),1,hour_func)
```
```{r,echo=FALSE,message=FALSE}
# Select data from certain periods
logi = capmetro$month>=9 &  capmetro$month<=11
data = capmetro[logi,]
```

We summarize the data and calculate average boardings
```{r,echo=FALSE,message=FALSE}
# Calculate average boardings
data_summary = data %>%
  group_by(hour,month,day) %>%
  summarize(mean_boarding=mean(boarding))
print(data_summary[1:18,])
```

\newpage
The faceted line plot is as follows.
```{r fig1,echo=FALSE,message=FALSE,fig.width=7.5,fig.height=9,fig.align = "center"}
data_summary$month = factor(data_summary$month)
p0 = ggplot(data=data_summary,aes(x=hour,y=mean_boarding,group=month,color=month))+ geom_line()+facet_wrap(~day, nrow=4)+ggtitle("Mean Boarding in different time intervals")+theme(plot.title = element_text(hjust = 0.5))
p0
```

\newpage
\textbf{(Q1\_1\_1) Does the hour of peak boardings change from day to day, or is it broadly similar across days?}

Based on the figure, It is broadly similar across days. At about hour 15.

\vspace{12pt}
\textbf{(Q1\_1\_2) Why do you think average boardings on Mondays in September look lower, compared to other days and months?}

September is the beginning of Fall semester, which means there is less people on campus. On Monday, perhaps there is less courses than other days.

\vspace{12pt}
\textbf{(Q1\_1\_3) Similarly, why do you think average boardings on Weds/Thurs/Fri in November look lower?}

There are many midterm exams in November, which means students stay in the dorm to study for the exams without having to go outside.

\newpage
\subsection{scatter plots}
```{r,echo=FALSE,message=FALSE}
#We first desigin indicator for weekdays/weekend.
# is_weekend 1 if it's a weekend
data$is_weekend = as.numeric(data$day=="Saturday" | data$day=="Sunday")
transform_func = function(num){
  if (num==1){
    return("weekend")
  }
  else{
    return("weekday")
  }
}
data$is_weekend_indicator = apply(array(data$is_weekend),1,transform_func)
```

The figure is as follows.
```{r fig2,echo=FALSE,message=FALSE,fig.width=7.5,fig.height=8,fig.align = "center"}
p0 = ggplot(data=data,aes(x=temperature,y=boarding,color=is_weekend_indicator))+
geom_point(alpha=0.3)+facet_wrap(~hour, nrow=4)+
ggtitle("The Relationship between Boarding and Temperature in different Hours")+
theme(plot.title = element_text(hjust = 0.5))
p0
```

\newpage
\textbf{(Q1\_2\_1) When we hold hour of day and weekend status constant, does temperature seem to have a noticeable effect on the number of UT students riding the bus?}

Just from the plot above, temperature doesn't have a noticeable effect on the number of UT students riding the bus.


\newpage
\section{Problem 2: Saratoga house prices}
\subsection{The Best Linear Model}
```{r,echo=FALSE,message=FALSE}
#We first load data.
library(tidyverse)
library(ggplot2)
library(rsample)
library(mosaic)
library(ModelMetrics)
data(SaratogaHouses)
```
The average of 5-fold corss-validation Rmse is used to evaluate a certain model.
```{r,echo=FALSE,message=FALSE,warning=FALSE}
# Metric definition(Use a number to evaluate the performance of a certain model),rmse is adopted.
library(caret)
evaluate_func = function(my_str,data){
  set.seed(100)
  indexs = createFolds(1:dim(data)[1], k = 5, list = TRUE, returnTrain = FALSE)
  err_result = c()
  for (i in 1:5){
    test_data = data[indexs[[i]],]
    train_data = data[-indexs[[i]],]
    lm_model = lm(as.formula(my_str), data=train_data)
    err = rmse(lm_model, test_data)
    err_result = c(err_result,err)
  }
  return(mean(err_result))
}

middle_str = "price ~ lotSize + age + livingArea + bedrooms +
fireplaces + bathrooms + rooms + heating + fuel + centralAir"
#print(evaluate_func(middle_str,SaratogaHouses))
```

The cross-validation Rmse of middle model is 65989.29. Our target is very simple, to find a model with cross-validation rmse lower than 65989.29. A greedy algorithm is used for feature selection, and the results are as follows.
```{r,echo=FALSE,message=FALSE}
x_names = colnames(SaratogaHouses)[2:length(colnames(SaratogaHouses))]
y_name = "price"
count = 0
rmse_record = Inf
best_name = ""
best_test_str = ""
best_record_previous = Inf
result_collect = c()

while(TRUE){
  for (name in x_names){
    if (count == 0){
     test_str = paste(y_name,"~",name,sep="")
    }
    else{
     test_str = paste(y_name,"+",name,sep="")
    }
    err = evaluate_func(test_str,SaratogaHouses)
    if (err<rmse_record){
      rmse_record = err
      best_name = name
      best_test_str = test_str
    }
  }
  if (rmse_record<best_record_previous){
    best_record_previous = rmse_record
  }
  else{
    break
  }
  y_name = best_test_str
  result_collect = c(result_collect,c(y_name,rmse_record))
  count = count + 1
  x_names = setdiff(x_names,best_name)
  if (length(x_names)==0){
    break
  }
}
print(result_collect)
```
```{r,echo=FALSE,message=FALSE}
best_str = "price~livingArea+landValue+bathrooms+waterfront+newConstruction+
heating+lotSize+centralAir+age+rooms+bedrooms+fuel+pctCollege+sewer+fireplaces"
print(evaluate_func(best_str,SaratogaHouses))
```

The best variables are:
```{r,echo=TRUE,message=FALSE}
# "price~livingArea+landValue+bathrooms+waterfront+newConstruction+
# heating+lotSize+centralAir+age+rooms+bedrooms+fuel+pctCollege+sewer+fireplaces"
# Error:57828.05
```

Corresponding cross-validation error is 57828.05, lower than 65989.29 from the medium model.

In all, we successfully overperform the medium model!.

\newpage
\subsection{The Best KNN}

Package kknn is used, and we slightly modify the evaluation function.
```{r,echo=FALSE,message=FALSE}
library(kknn)
knn_evaluate_func = function(my_str,data){
  set.seed(100)
  indexs = createFolds(1:dim(data)[1], k = 5, list = TRUE, returnTrain = FALSE)
  err_result = c()
  for (i in 1:5){
    test_data = data[indexs[[i]],]
    train_data = data[-indexs[[i]],]
    model = train.kknn(as.formula(my_str),train_data,scale = TRUE)
    test_prediction = predict(model,test_data)
    test_actual = test_data$price
    err = rmse(actual=test_actual, predicted=test_prediction)
    err_result = c(err_result,err)
  }
  return(mean(err_result))
}
```

Greedy algorithm is adopted again to select the best feature combination, the results are as follows.
```{r,echo=FALSE,message=FALSE}
x_names = colnames(SaratogaHouses)[2:length(colnames(SaratogaHouses))]
y_name = "price"
count = 0
rmse_record = Inf
best_name = ""
best_test_str = ""
best_record_previous = Inf
result_collect = c()

while(TRUE){
  for (name in x_names){
    if (count == 0){
     test_str = paste(y_name,"~",name,sep="")
    }
    else{
     test_str = paste(y_name,"+",name,sep="")
    }
    err = knn_evaluate_func(test_str,SaratogaHouses)
    if (err<rmse_record){
      rmse_record = err
      best_name = name
      best_test_str = test_str
    }
  }
  if (rmse_record<best_record_previous){
    best_record_previous = rmse_record
  }
  else{
    break
  }
  y_name = best_test_str
  result_collect = c(result_collect,c(y_name,rmse_record))
  count = count + 1
  x_names = setdiff(x_names,best_name)
  if (length(x_names)==0){
    break
  }
}
print(result_collect)
```

\newpage
\subsection{Analysis}

The best variables and cross-validation error for KNN is
```{r,echo=TRUE,message=FALSE}
# "price~livingArea+landValue+age+pctCollege+waterfront+newConstruction"
# 58061.7316651973
```

The best variables and cross-validation error for linear model is
```{r,echo=TRUE,message=FALSE}
# "price~livingArea+landValue+bathrooms+waterfront+newConstruction+heating+
# lotSize+centralAir+age+rooms+bedrooms+fuel+pctCollege+sewer+fireplaces"
# 57828.05
```

Although the cross-validation error is lower for linear model, I still believe knn
is better, as it uses only 6 variables to achieve its lowest error.

Moreover,$\frac{58061.7-57828.05}{57828.05}=0.00404$, not very much.

\newpage
\section{Problem 3: Classification and retrospective sampling}

Summary of the data.
```{r,echo=FALSE,message=FALSE}
library(vcd)
library(ggplot2)
library(tidyverse)

german_credit <- read.csv("data/german_credit.csv", stringsAsFactors=TRUE)
summary(german_credit)
# Default：300 Not Default: 700
```

Count of the data.
```{r,echo=FALSE,message=FALSE}
counts <- table(german_credit$history)
counts
# good: 89 poor:618 terrible: 293
```

\newpage
For categories poor and terrible we see the number of nodefault are greater than the number of default. For categories good we see the number of default is greater than the number of nodefault.
```{r fig 4,echo=FALSE,message=FALSE,fig.width=4,fig.height=4,fig.align = "center"}
ggplot(german_credit, aes(history, ..count..)) +
  geom_bar(aes(fill = as.factor(Default)), position = "dodge")
```

Train and test split.
```{r,echo=FALSE,message=FALSE}
# sampling
set.seed(1234)
train <- sample(nrow(german_credit),0.7*nrow(german_credit))
german_credittrain <- german_credit[train,]
german_credittest <- german_credit[-train,]
print("train_summary")
table(german_credittrain$history)
print("test_summary")
table(german_credittest$history)
```

\newpage
\textbf{(Q3\_1) Make a bar plot of default probability by credit history}

```{r fig 5,echo=FALSE,message=FALSE,fig.width=4.5,fig.height=5,fig.align = "center"}
credit.glm1 <- glm(Default ~ history, family = binomial, german_credittrain )
DefaultProppoor <-  2.718281828459^(0.528-1.261*1) / (1+2.718281828459^(0.528-1.261*1))
DefaultPropterrible <- 2.718281828459^(0.528-2.241*1) / (1+2.718281828459^(0.528-2.241*1))
DefaultPropgood <- 1-(DefaultProppoor+DefaultPropterrible)
DefaultProp<- c(DefaultPropterrible,DefaultProppoor,DefaultPropgood)
history1<- c("terrible", "poor", "good")
Defaulthistory<- data.frame(history1,DefaultProp)
ggplot(Defaulthistory) +
  geom_bar(stat='identity',aes(x=history1, y=DefaultProp), color='blue', alpha=0.1) +
  labs(title="default probability by credit history", x="history1", y="DefaultProp")+
theme(plot.title = element_text(hjust = 0.5))
```

\newpage
\textbf{(Q3\_2) Build a logistic regression model for predicting default probability, using the variables duration + amount + installment + age + history + purpose + foreign.}

The summary of model is as follows.

```{r,echo=FALSE,message=FALSE}
fit.logit <- glm( Default ~ duration + amount + installment + age + history + purpose + foreign, family = binomial(), data=german_credittrain)
summary(fit.logit)
coef(fit.logit)
coef(fit.logit) %>% round (3)
exp(coef(fit.logit))
```

\newpage
\textbf{(Q3\_3) What do you notice about the history variable vis-a-vis predicting defaults? What do you think is going on here?}

According to the graph, the default probablitity will become higher as the borrower's credit rating is better.Because the bank matched each default with similar sets of loans that had not defaulted, including all reasonably close matches in the analysis.The sample of People with good credit is too small to lower the accuracy and they usually have fewer default samples. This resulted in a substantial oversampling of defaults


We use the confusion matrix to check out-of-sample performance
```{r,echo=FALSE,message=FALSE}
prob <- predict(fit.logit, german_credittest, type="response")
logit.pred <- factor(prob > 0.5, levels=c(FALSE, TRUE),
                    labels=c("Nodefault", "default"))
logit.perf <- table(german_credittest$Default, logit.pred,
                    dnn=c("Actual", "Predicted"))
logit.perf
```
accuracy= (188+22)/300= 0.70

An example for predict default history by using the logistic model
```{r,echo=FALSE,message=FALSE}
german_credittest[1,]
predictions_1st <- predict(fit.logit, newdata = german_credittest[1,], type = "response")
predictions_1st
```

We could see that the defaulting probability for the 1st player in the test set is about 45.77%.


\newpage
\textbf{(Q3\_4) In light of what you see here, do you think this data set is appropriate for building a predictive model of defaults, if the purpose of the model is to screen prospective borrowers to classify them into "high" versus "low" probability of default? Why or why not---and if not, would you recommend any changes to the bank's sampling scheme?}

I think this data set is not appropriate for building a predictive model for defaults. Because the bank attempted to match each default with similar sets of loans that had not defaulted, including all reasonably close matches in the analysis. This resulted in a substantial oversampling of defaults, relative to a random sample of loans in the bank's overall portfolio.

The bank’s sample size should be as large as possible. The more closer to the overall conditions, the more accurate the predictive outcome is.


\newpage
\section{Problem 4: Children and hotel reservations}
```{r,echo=FALSE,message=FALSE,warning=FALSE}
library(tidyverse)
library(ggplot2)
library(ModelMetrics)
library(caret)
library(lubridate)
```

\subsection{Model building}

\textbf{(Q4\_1)Compare the out-of-sample performance of the following models}

We first load the data.
```{r,echo=FALSE,message=FALSE,warning=FALSE}
library(plotROC)
hotelsdev = read.csv("data/hotels_dev.csv",stringsAsFactors=TRUE)
```

The mean f1-score from 5-fold cross-validation is applied to evaluate the performance of certain model.
```{r,echo=FALSE,message=FALSE,warning=FALSE}
evaluate_func = function(my_str,data){
  set.seed(100)
  indexs = createFolds(1:dim(data)[1], k = 5, list = TRUE, returnTrain = FALSE)
  err_result = c()
  for (i in 1:5){
    test_data = data[indexs[[i]],]
    train_data = data[-indexs[[i]],]
    model = glm(as.formula(my_str), data=train_data,family = "binomial")
    truth = test_data$children
    prediction = predict(model,newdata = test_data,type = "response")
    err_result = c(err_result,f1Score(truth,prediction))
  }
  return(mean(err_result))
}
```

\textbf{1 Baseline model 1}

The summary of Baseline 1(fitting on the whole data set):
```{r,echo=FALSE,message=FALSE,warning=FALSE}
baseline_1_str = "children ~ market_segment + adults + customer_type + is_repeated_guest"
model = glm(as.formula(baseline_1_str),data=hotelsdev,family = "binomial")
prediction = predict(model,type = "response")
truth = hotelsdev$children
summary(model)
```

The f1-score of fitting
```{r,echo=FALSE,message=FALSE,warning=FALSE}
f1Score(truth,prediction)
```
The 5-fold cross-validation f1-score of Baseline 1:
```{r,echo=FALSE,message=FALSE,warning=FALSE}
print(evaluate_func(baseline_1_str,hotelsdev))
```

\newpage
\textbf{2 Baseline model 2}

The summary of Baseline 2(fitting on the whole data set):
```{r,echo=FALSE,message=FALSE,warning=FALSE}
baseline_2_str = "children ~ (. - arrival_date)"
model = glm(as.formula(baseline_2_str),data=hotelsdev,family = "binomial")
prediction = predict(model,type = "response")
truth = hotelsdev$children
summary(model)
```

The f1-score of fitting
```{r,echo=FALSE,message=FALSE,warning=FALSE}
f1Score(truth,prediction)
```
The 5-fold cross-validation f1-score of Baseline 2:
```{r,echo=FALSE,message=FALSE,warning=FALSE}
print(evaluate_func(baseline_2_str,hotelsdev))
```

\vspace{12pt}
\textbf{3 Best linear model}

We generate the time-stamp of year, month, day, and day of week from arrival_date.
```{r,echo=FALSE,message=FALSE,warning=FALSE}
hotelsdev = mutate(hotelsdev,
                   arrival_date = ymd(arrival_date))
hotelsdev = mutate(hotelsdev, 
                   wday = wday(arrival_date) %>% factor(), 
                   day = day(arrival_date) %>% factor(),
                   month = month(arrival_date) %>% factor(),
                   year = year(arrival_date))
hotelsdev$arrival_date = as.character(hotelsdev$arrival_date)
hotelsdev <- subset(hotelsdev, select = -c(arrival_date)) 
```

Then, we use greedy algorithm to find the best feature combination.
```{r,echo=FALSE,message=FALSE,warning=FALSE}
# x_names = colnames(hotelsdev)[-6]
# y_name = "children"
# count = 0
# rmse_record = 0
# best_name = ""
# best_test_str = ""
# best_record_previous = 0
# result_collect = c()
# 
# while(TRUE){
#   for (name in x_names){
#     if (count == 0){
#       test_str = paste(y_name,"~",name,sep="")
#     }
#     else{
#       test_str = paste(y_name,"+",name,sep="")
#     }
#     err = evaluate_func(test_str,hotelsdev)
#     if (err>rmse_record){
#       rmse_record = err
#       best_name = name
#       best_test_str = test_str
#     }
#   }
#   if (rmse_record>best_record_previous){
#     best_record_previous = rmse_record
#   }
#   else{
#     break
#   }
#   y_name = best_test_str
#   result_collect = c(result_collect,c(y_name,rmse_record))
#   if (rmse_record>0.505){
#     break
#   }
#   print(c(y_name,rmse_record))
#   count = count + 1
#   x_names = setdiff(x_names,best_name)
#   if (length(x_names)==0){
#     break
#   }
# }
print("children~reserved_room_type")                                          
print("0.364107552305935")                                                    
print("children~reserved_room_type+hotel")                                     
print("0.506343075649843")                                                    
print("children~reserved_room_type+hotel+previous_cancellations")              
print("0.506437707712283")                                                     
print("children~reserved_room_type+hotel+previous_cancellations+booking_changes")  
print("0.506463838208211")  
```

The best feature combination is (cross-validation f1-score 0.50646)
```{r,echo=FALSE,message=FALSE,warning=FALSE}
print("children~reserved_room_type+hotel+previous_cancellations+booking_changes")  
```

The summary of best model:
```{r,echo=FALSE,message=FALSE,warning=FALSE}
best_str = "children~reserved_room_type+hotel+previous_cancellations+booking_changes"
model = glm(as.formula(best_str),data=hotelsdev,family = "binomial")
summary(model)
prediction = predict(model,type = "response")
truth = hotelsdev$children
```

The f1-score of fitting
```{r,echo=FALSE,message=FALSE,warning=FALSE}
f1Score(truth,prediction)
```

The 5-fold cross-validation f1-score of best model:
```{r,echo=FALSE,message=FALSE,warning=FALSE}
print(evaluate_func(best_str,hotelsdev))
```

\vspace{12pt}
\textbf{4 Analysis}

The cross-validation f1-scores of 3 models are as follows:

Baseline1 model 1: 0

Baseline1 model 2: 0.4642258

Best model: 0.5064638

The best model is the best model with the highest f1-score.

\newpage
\subsection{Model validation: step 1}

\textbf{(Q4\_2)Produce an ROC curve for your best model, using the data in hotels_val: that is, plot TPR(t) versus FPR(t) as you vary the classification threshold t.}

We first fit on the hotelsdev data. Then we load hotels_val to conduct validation and draw ROC graph.
```{r,echo=FALSE,message=FALSE,warning=FALSE}
best_str = "children~reserved_room_type+hotel+previous_cancellations+booking_changes"
model = glm(as.formula(best_str),data=hotelsdev,family = "binomial")
```
```{r,echo=FALSE,message=FALSE,warning=FALSE}
# Then we load hotels_val to conduct validation and draw ROC graph.
hotelsval = read.csv("data/hotels_val.csv")
prediction = predict(model,newdata = hotelsval,type = "response")
truth = hotelsval$children
roc_data = cbind(truth,prediction)
roc_data = data.frame(roc_data)
basicplot <- ggplot(roc_data, aes(d = truth,m = prediction))+ geom_roc() 
```

The plot is as follows:
```{r fig21,echo=FALSE,message=FALSE,fig.width=3.5,fig.height=4,fig.align = "center"}
advanced_plot = basicplot + 
  style_roc(theme = theme_grey) +
  theme(axis.text = element_text(colour = "blue"),plot.title = element_text(hjust = 0.5))+ 
  annotate("text", x = .75, y = .25,label = paste("AUC =", round(calc_auc(basicplot)$AUC, 2))) +
  scale_x_continuous("1 - Specificity", breaks = seq(0, 1, by = .1))
advanced_plot
```

\newpage
\subsection{Model validation: step 2}

\textbf{(Q4\_3)How well does your model do at predicting the total number of bookings with children in a group of 250 bookings? Summarize this performance across all 20 folds of the val set in an appropriate figure or table.}

We first fit on the hotelsdev data. Then we load hotels_val to calculate prediction accuracy.
```{r,echo=FALSE,message=FALSE,warning=FALSE}
best_str = "children~reserved_room_type+hotel+previous_cancellations+booking_changes"
model = glm(as.formula(best_str),data=hotelsdev,family = "binomial")
# Then we load hotels_val to conduct validation and draw ROC graph.
hotelsval = read.csv("data/hotels_val.csv")
```
The hotels_val is divided into 20 folds.
```{r,echo=FALSE,message=FALSE,warning=FALSE}
set.seed(100)
indexs = createFolds(1:dim(hotelsval)[1], k = 20, list = TRUE, returnTrain = FALSE)
err_result = c()
for (i in 1:20){
  slice_data = hotelsval[indexs[[i]],]
  prediction = predict(model,newdata = slice_data,type = "response")
  predict_num = round(mean(prediction)*dim(slice_data)[1])
  ### expected number of bookings with children for that fold.
  actual_num = sum(slice_data$children)
  err_result = rbind(err_result,c(actual_num,predict_num))
} 
```
The following is the summary of expected number of bookings with children for that fold and actual number for that fold.
```{r,echo=FALSE,message=FALSE,warning=FALSE}
err_result = data.frame(err_result)
colnames(err_result)=c("actual_num","predict_num") 
err_result$difference = err_result$predict_num-err_result$actual_num
print(err_result)
```

The following figure demonstrates the distribution of difference beween actual number and expected number among 20 folds.
```{r fig22,echo=FALSE,message=FALSE,fig.width=2.5,fig.height=3,fig.align = "center"}
p0 = ggplot(data=err_result) + 
  geom_histogram(aes(x=difference)) 
p0
```






